Is the Future of Authenticity In Our Heads?\\\medskip \large Moving Passthoughts From the Lab to the World

### 2017-07-03 paper.tex
Draft to shepard

% Created 2017-07-03 Mon 10:49
% Intended LaTeX compiler: pdflatex
\documentclass[sigconf]{acmart}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{grffile}
\usepackage{longtable}
\usepackage{wrapfig}
\usepackage{rotating}
\usepackage[normalem]{ulem}
\usepackage{amsmath}
\usepackage{textcomp}
\usepackage{amssymb}
\usepackage{capt-of}
\usepackage{hyperref}
\input{authors}
\setcopyright{rightsretained}
\acmDOI{10.475/123_4}
\acmISBN{123-4567-24-567/08/06}
\acmConference[NSPW '17]{New Security Paradigms Workshop}{October 2017}{Islamorada, Florida, USA}
\acmYear{2017}
\copyrightyear{2017}
\acmPrice{15.00}
\date{\today}
\title{Is the Future of Authenticity In Our Heads?\\\medskip
\large Moving Passthoughts From the Lab to the World}
\hypersetup{
 pdfauthor={},
 pdftitle={Is the Future of Authenticity In Our Heads?},
 pdfkeywords={},
 pdfsubject={},
 pdfcreator={Emacs 25.1.1 (Org mode 9.0.4)}, 
 pdflang={English}}
\begin{document}

@@

\section{Introduction}
\label{sec:orgf1c0eda}

@@

\section{Background}
\label{sec:org6b9dd4a}

@@

\subsection{One-step, multi-factor authentication}
\label{sec:org526fbba}

@@

\subsection{Passthought authentication}
\label{sec:org6b3c69e}

@@

\subsection{Passthoughts using in-ear EEG}
\label{sec:org0f217d8}

@@

\section{User attitudes and perceptions}
\label{sec:org0009958}

@@

\subsection{Contending with mind-reading machines}
\label{sec:org5b8dcbd}

@@

\subsection{What (do you think) EEG can reveal about a person?}
\label{sec:orga19eaac}

@@

\section{Diversity and security of passthoughts}
\label{sec:org4fff047}

@@

A more subtle, but closely related question surrounds how passthought EEG recordings relate statistically to non-passthought EEG recordings.
In other words, we do not know how the particular passthoughts observed in past work are drawn from the distribution of EEG signals that an individual produces over the course of their day.
This blind-spot poses a possible challenge to passthought's vulnerability to dictionary-style cracking.
If an attacker has a large enough corpus of EEG readings, do some passthoughts start to look as guessable as \emph{password1234}?
By answering such questions, we could design data-driven policies for, e.g., how many retry attempts passthought authenticators should allow.

\section{Two studies on passthoughts}
\label{sec:org9bd76c2}

@@

\subsection{A real-time passthought authenticator}
\label{sec:org65e0cf5}

@@

\subsubsection{Study protocol}
\label{sec:org0bd4853}

@@

\subsubsection{The effect of feedback}
\label{sec:org4f731b0}

@@

\subsubsection{Exploring continuous re-calibration}
\label{sec:orgfa27859}

@@

\subsection{A longitudinal study on brainwave monitoring}
\label{sec:org7ef7168}

@@

Unfortunately, these challenges (particularly those around shifting neural signals) also make it difficult to produce a passthought authenticator that works with any reliability in real-life contexts.
This makes a longitudinal study with a working authenticator impractical for the time being.
However, we may still perform a longitudinal study that allows us to interrogate the usability aspects around (and attitudes about) passthoughts specifically, and EEG generally.
In so doing, we may also collect a larger and more diverse corpus of passthoughts, which can be used to address the paucity of data we face today.
A technology probe or diary study \cite{Gaver1999} could help address both of these issues at the same time.

Of course, this study would be no substitute for a working, online passthoughts authentication system.
Instead, this study aims to collect useful data before such a system exists.
It will not only elicit beliefs, 
but also allow us to collect larger datasets, 
and to catch technical issues in sensing devices and collection platforms.

\subsubsection{Study protocol}
\label{sec:org7967cd6}

@@

\subsubsection{A more diverse corpus}
\label{sec:org6c1e1e9}

@@

\subsubsection{The space of possible passthoughts}
\label{sec:org9ef3205}
In another potentially fruitful analysis, such a corpus will allow us to perform statistical analysis of how passthoughts are drawn from the overall distribution of EEG signals. 
Using multi-dimensional clustering algorithms such as t-SNE \cite{VanDerMaaten2008} 
could assist us in understanding how particular passthoughts relate to other EEG signals that an individual expresses involuntarily throughout the day. 
These clusters will help us understand how likely or unlikely we are to observe a given passthought in context of a particular person's neural signals
Such analysis between subjects could help shed light how given passthoughts are expressed uniquely between individuals.

@@

\subsubsection{Usability and attitudes}
\label{sec:orge398ca0}

@@

\section{Privacy, Security: Choices, Tradeoffs}
\label{sec:orgbb7cf43}

@@

\subsection{Privacy}
\label{sec:orga1e2eb9}
As of yet, it is still not well understood what EEG signals might reveal about a person.
EEG signals that are not anonymized could come to be seen as private in the face of new methods of analysis.
(If your brainwaves can authenticate you, could they also uniquely identify you, even if your name is redacted?)
Differential privacy \cite{Dwork2014} presents one approach to dealing with the risk of privacy breaches with EEG signals.
By adding noise to datasets, differentially private databases can make strong guarantees about the likelihood of a de-anonymization attack on particular database queries.

\subsection{Security}
\label{sec:org759dea8}
Device security presents another risk to passthought authentication.
Since EEG devices will transmit data, likely wirelessly \cite{Mihajlovic2015}, their data may be intercepted, depending on the security properties of the underlying transit protocol. 
When transferring authentication credentials in passthoughts, the ability to snoop on authentication attempts could present a dangerous attack vector.

@@

\subsection{Tradeoffs between security and privacy}
\label{sec:orgd71b4f9}

@@

\section{Further Future Directions}
\label{sec:org9fefd9f}

@@

\subsection{Continuous authentication}
\label{sec:org40ff196}

@@

\subsection{Organic passwords}
\label{sec:org74d0a56}

If EEG signals are nonstationary (changing over time), passthoughts will require continuous re-calibration to maintain decent accuracy \cite{Vidaurre2006a}.
This feature of BCIs could have an unexpected benefit to security. 
If an individual's expression of their passthought in EEG is always changing, 
passthoughts themselves are effectively evergreen, automatically replaced or updated by nature of the authentication paradigm.
This feature could improve security, as an attacker able to compromise a passthought's EEG signature may not be able to log into the system in a few weeks time,
unless they are able to realistically mutate the signal over authentication attempts.
This feature of EEG also gives passthoughts a possible advantage over other methods for behavioral authentication, such as voice or keystroke dynamics \cite{Monrose1997}, which may change more slowly for individuals, if they change at all.
Future work should investigate this claim, perhaps using a longitudinal corpus such as the one described above.
\subsection{Authentication and the self}
\label{sec:org740e10b}

@@

Such a scenario raises serious ethical, legal, and even philosophical questions. 
How does such a system conform to accepted definitions of a ``person''?
Who is a person to make decisions for their future self?
What are possible vectors for abuse?
In any case, this property of an authentication is, as far as we are aware, novel, 
and should be considered as we learn more about the strengths, weaknesses, and particular affordances of this developing method for authentication.
\subsection{Passthoughts by any other sensor?}
\label{sec:org8dc3b22}

@@

\subsection{Health, neuroscience and BCIs}
\label{sec:org037222f}

@@

\section{Conclusion}
\label{sec:orgc0cdb64}

>>>>

% Created 2017-07-03 Mon 12:23
% Intended LaTeX compiler: pdflatex
\documentclass[sigconf]{acmart}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{grffile}
\usepackage{longtable}
\usepackage{wrapfig}
\usepackage{rotating}
\usepackage[normalem]{ulem}
\usepackage{amsmath}
\usepackage{textcomp}
\usepackage{amssymb}
\usepackage{capt-of}
\usepackage{hyperref}
\input{authors}
\setcopyright{rightsretained}
\acmDOI{10.475/123_4}
\acmISBN{123-4567-24-567/08/06}
\acmConference[NSPW '17]{New Security Paradigms Workshop}{October 2017}{Islamorada, Florida, USA}
\acmYear{2017}
\copyrightyear{2017}
\acmPrice{15.00}
\date{\today}
\title{Is the Future of Authenticity In Our Heads?\\\medskip
\large Moving Passthoughts From the Lab to the World}
\hypersetup{
 pdfauthor={},
 pdftitle={Is the Future of Authenticity In Our Heads?},
 pdfkeywords={},
 pdfsubject={},
 pdfcreator={Emacs 25.1.1 (Org mode 9.0.4)}, 
 pdflang={English}}
\begin{document}

@@

\section{Introduction}
\label{sec:orge209717}

@@

\section{Background}
\label{sec:org6827efb}

@@

\subsection{One-step, multi-factor authentication}
\label{sec:orgffe92dc}

@@

\subsection{Passthought authentication}
\label{sec:org93d707a}

@@

\subsection{Passthoughts using in-ear EEG}
\label{sec:org09042b7}

@@

\section{User attitudes and perceptions}
\label{sec:orgdbbbb66}

@@

\subsection{Contending with mind-reading machines}
\label{sec:org45e1492}

@@

\subsection{What (do you think) EEG can reveal about a person?}
\label{sec:org6df83c3}

@@

Our qualitative data revealed that subjects in both groups generally believed EEG to reveal various details about the mind, mood, emotions, and identity.
We asked subjects to reflect on why they answered the way they did during the ranking task (Figure \ref{fig:rank}).
In the Health-e-Heart group, several subjects gave relatively specific explanations as to why they ranked EEG hihgly.

\begin{quote}
\emph{(S24) I assume some information can be gleaned from brain wave activity in various parts of the brain related to rewards or executive control, but without accompanying information, it may be difficult to discover my thoughts.}
\end{quote}

\begin{quote}
\emph{(S23) EEGs note parts of the brain that are active.  Again, in conjunction with other measurements, I suspect that some sense of what one is thinking and feeling could be learned.}
\end{quote}

\begin{quote}
\emph{(S91) I would rate this relatively high on the list because science has shown that we can detect a lot about which areas of the brain are accessed and at which times.  This can tell a person a lot about what they might be thinking and especially how they are feeling.}
\end{quote}

While these explanations range somewhat in their specificity and confidence,
they share the general sentiment that EEGs can be revealing. Subjects in the Mechanical Turk condition broadly shared this belief, though tended to use less physiological detail in their explanations.

\begin{quote}
\emph{(S157) Brain activity can pinpoint exact emotions by monitoring certain areas on the brain.}
\end{quote}

\begin{quote}
\emph{(S130) Brainwaves could tell you a lot more about what someone is thinking and feeling. You could measure the patterns of brainwaves in an experiment.}
\end{quote}

Meanwhile, some subjects from both groups did not fit this trend. Ten subjects ranked EEG low in its ability to measure what a person is thinking or feeling. Their qualitative answers revealed a diverse set of reasons for this ranking.
Three subjects indicated a general lack of faith in brainwave's reliabilty.

\begin{quote}
\emph{(S20) I don't think we have the ability to translate brainwaves into thoughts or emotions.}
\end{quote}

\begin{quote}
\emph{(S101) EEG is very nonspecific and rarely can tell details reliably.}
\end{quote}

\begin{quote}
\emph{(S138) Possible but not accurate.}
\end{quote}

These explanations broadly centered around EEG as a signal.
They range somewhat in their confidence, from a fundamental skepticism (S20) to caveats about possible accuracy or specificity (S101, S138).
In contrast to these three subjects, S10 ranked EEG low because s/he
felt the premise of a consumer grade EEG was implausible.

\begin{quote}
\emph{(S10) I assume that scientists can identify by brain patterns what others are feeling and thinking based off of years of research.  I've never heard of a consumer grade eeg - and doubt it could be as powerful as a laboratory eeg.  If it is then I would be interested in this product.}
\end{quote}

This subject's explanation surfaces the practical differences in attitudes that people might have to a technology's theoretical existence,
and its realized existence as a consumer device. Future work could look more closely at 
how the presumed scientific authority of a brainscanning apparatus affects people's willingness to accept specific BCI applications such as passthoughts \cite{Ali2014a}.
Finally, one subject's skepticism what brainwaves can reveal stemmed from his/her personal medical experiences.

\begin{quote}
\emph{(S116) My son has absence seizures, so his brainwaves change.}
\end{quote}

This quote highlights how individuals' life experiences
might shape the way they engage (or refuse to engage) with brain-sensing devices.
In general, this quote and others motivate the need for a rich, qualitative understanding of people's first-hand experiences with brainscanning devices,
as well as data collection,
in order to understand what role BCI applications such as passthoughts could play in day-to-day life.
\section{Diversity and security of passthoughts}
\label{sec:org3330c22}

@@

A more subtle, but related question surrounds how passthought EEG recordings compare to non-passthought EEG recordings.
In other words, we do not know how the particular passthoughts observed in past work are drawn from the distribution of EEG signals that an individual produces over the course of their day.
This blind-spot poses a possible challenge to passthought's vulnerability to dictionary-style cracking.
If an attacker has a large enough corpus of EEG readings, do some passthoughts start to look as guessable as \emph{password1234}?
By answering such questions, we could design data-driven policies for, e.g., how many retry attempts passthought authenticators should allow.

\section{Two studies on passthoughts}
\label{sec:org7ca404f}

@@

\subsection{A real-time passthought authenticator}
\label{sec:orgf20bd64}

@@

\subsubsection{Study protocol}
\label{sec:orge48e6fe}

@@

\subsubsection{The effect of feedback}
\label{sec:orgf1fddaa}

@@

\subsubsection{Exploring continuous re-calibration}
\label{sec:orgef60fad}

@@

\subsection{A longitudinal study on brainwave monitoring}
\label{sec:orgbea5086}

@@

Unfortunately, these challenges (particularly those around shifting neural signals) also make it difficult to produce a passthought authenticator that works with any reliability in real-life contexts.
This makes a longitudinal study with a working authenticator impractical for the time being.
However, we may still perform a longitudinal study that allows us to interrogate the usability aspects around (and attitudes about) passthoughts specifically, and EEG generally.
In so doing, we may also collect a larger and more diverse corpus of passthoughts, which can be used to address the paucity of data we face today.
This section describes a technology probe \cite{Gaver1999} that could help address both of these issues at the same time.

\subsubsection{Study protocol}
\label{sec:org9be49e3}

@@

\subsubsection{A more diverse corpus}
\label{sec:org4f6869d}

@@

\subsubsection{The space of possible passthoughts}
\label{sec:orgb19bfc8}
In another potentially fruitful analysis, such a corpus will allow us to perform statistical analysis of how passthoughts are drawn from the overall distribution of EEG signals. 
Using multi-dimensional clustering algorithms such as t-SNE \cite{VanDerMaaten2008} 
could assist us in understanding how particular passthoughts relate to other EEG signals that an individual expresses involuntarily throughout the day. 
These clusters will help us understand how likely or unlikely we are to observe a given passthought in context of a particular person's neural signals
Such analysis between subjects could help shed light how given passthoughts are expressed uniquely between individuals.

@@

\subsubsection{Usability and attitudes}
\label{sec:org488cdca}

@@

\subsubsection{Limitations}
\label{sec:org794d9d1}

This study would be no substitute for a working, online passthoughts authentication system.
Instead, this study aims to collect useful data before such a system exists.
It will not only elicit beliefs, 
but also allow us to collect larger datasets, 
and to catch technical issues in sensing devices and collection platforms.

Even in this goal, the proposed study has a few limitations.
First, it is unclear how closely the study protocol maps to actual passthought use in the wild.
For example, people who use passthoughts may not wish to wear an EEG device all day, as our subjects would.
Furthermore, the system proposed here does not provide a realistic authentication context, in that subjects
are asked to use the system at pre-defined points during the day. 
Future work could create more realistic concepts, perhaps ones in which subjects have an intrinsic motivation (or stake) for the authenticator to work correctly.

\section{Privacy, Security: Choices, Tradeoffs}
\label{sec:org12abefb}

@@

\subsection{Privacy}
\label{sec:org13854dd}
As of yet, it is still not well understood what EEG signals might reveal about a person.
EEG signals that are not anonymized could come to be seen as private in the face of new methods of analysis.
(If your brainwaves can authenticate you, could they also uniquely identify you, even if your name is redacted?)
Differential privacy \cite{Dwork2014} presents one approach to dealing with the risk of privacy breaches with EEG signals.
By adding noise to datasets, differentially private databases can make strong guarantees about the likelihood of a de-anonymization attack on particular database queries.

\subsection{Security}
\label{sec:orgac01f42}
Device security presents another risk to passthought authentication.
Since EEG devices will transmit data, likely wirelessly \cite{Mihajlovic2015}, their data may be intercepted, depending on the security properties of the underlying transit protocol. 
When transferring authentication credentials in passthoughts, the ability to snoop on authentication attempts could present a dangerous attack vector.

@@

\subsection{Tradeoffs between security and privacy}
\label{sec:org3cb1742}

@@

\section{Further Future Directions}
\label{sec:orged2d30f}

@@

\subsection{Continuous authentication}
\label{sec:org4bbe60c}

@@

\subsection{Organic passwords}
\label{sec:org4ccf763}

If EEG signals are nonstationary (changing over time), passthoughts will require continuous re-calibration to maintain decent accuracy \cite{Vidaurre2006a}.
This feature of BCIs could have an unexpected benefit to security. 
If an individual's expression of their passthought in EEG is always changing, 
passthoughts themselves are effectively evergreen, automatically replaced or updated by nature of the authentication paradigm.
This feature could improve security, as an attacker able to compromise a passthought's EEG signature may not be able to log into the system in a few weeks time,
unless they are able to realistically mutate the signal over authentication attempts.
This feature of EEG also gives passthoughts a possible advantage over other methods for behavioral authentication, such as voice or keystroke dynamics \cite{Monrose1997}, which may change more slowly for individuals, if they change at all.
Future work should investigate this claim, perhaps using a longitudinal corpus such as the one described above.
\subsection{Authentication and the self}
\label{sec:orgc34f1a2}

@@

Such a scenario raises serious ethical, legal, and even philosophical questions. 
How does such a system conform to accepted definitions of a ``person''?
Who is a person to make decisions for their future self?
What are possible vectors for abuse?
In any case, this property of an authentication is, as far as we are aware, novel, 
and should be considered as we learn more about the strengths, weaknesses, and particular affordances of this developing method for authentication.
\subsection{Passthoughts by any other sensor?}
\label{sec:org55fcab6}

@@

\subsection{Health, neuroscience and BCIs}
\label{sec:orga1d8b9b}

@@

\section{Conclusion}
\label{sec:org74bc9a2}
